State of the Practice of Lattice Boltzmann Method Software Development

### 2021-07-05 StateOfPractice/Peter-Notes/Report/Report.tex
incorporated full report feedback from SS into LBM report

\documentclass[12pt, notitlepage]{article}
\usepackage{amsmath, mathtools}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{xr}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{xfrac}
\usepackage{tabularx}
\usepackage{float}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{titling}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage{indentfirst}

@@

We analyze the state of the practice of software development of Lattice Boltzmann Methods software domain by quantitatively and qualitatively measuring, and comparing, 23 software packages along quality attributes. A methodology for assessing the state of the practice of software development in scientific computing software domains is presented. A domain analysis of the Lattice Boltzmann Methods software family is made, and a candidate software package list is identified and curated. The packages are assessed to answer software development related research questions to understand how software quality is impacted by software development choices, including principles, processes, and tools. Software developers are interviewed to identify development pain points, and to identify how software quality is ensured. Quantitative data is used to rank the software packages using an analytical hierarchy process. The ranking designations are compared with rankings from the software development community. Recommendations for improving software along quality metrics are made. This knowledge can be used to guide future development of scientific computing software, specifically along quality attributes, and to reduce software quality failures.\\

@@

We analyze the development of software packages that use Lattice Boltzmann Methods (LBM), a class of computational fluid dynamics (CFD) methods, for fluid simulation. These LBM algorithms consider the behaviours of a collection of particles as a single unit of an intermediate size, and predict their positional probability as they move through a lattice structure. Figure \ref{circularflow} from \citep{chen2021phase} presents the streamlines of converged flow past a stationary circular cylinder with varied Reynolds number, a common task of LBM.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{circularflow}
		\caption{Streamlines of the flow past a stationary circular cylinder at Re = 10, 20, and 40}
		\label{circularflow}
	\end{center}
\end{figure}

@@

The packages are assessed for a given set of quality attributes using quantitative and qualitative data. The qualities are defined in Section \ref{softwarequalities}. Qualitative data is gathered from interviews with software package developers. The interview questions are listed in the Appendix in \ref{interviewquestions}. Quantitative data is measured using the quality metrics listed in the Appendix in \ref{measurementtemplate2}. In this context the purposes of the report are as follows:

@@

In the course of this assessment we updated the methodology that was used during previous assessments. Details of the new methodology are in \citep{methodology2020}. The previous set of research questions were critically assessed and modified. The updated questions are listed in Section \ref{purpose}. In this re-boot we collected more quantitative and qualitative data, focused on software quality attributes, and added more measures, including collecting empirical data and interviewing developers. We have also added a domain analysis to better characterize the functionality provided by the software, and have leveraged the expertise of a domain expert. As in past assessments, the collected data was combined to rank the software using the Analytic Hierarchy Process (AHP). The domain expert was consulted to verify the ordering.

@@

We analyze a filtered set of LBM software packages along quantitative and qualitative measures. Many of these measures are captured using surface measurements, which can be categorized as initial and easy to capture measurements, of the underlying quality; they may not represent the true quality of the software. Surface measurements are taken as they allow us to apply the same measurement, with reasonable effort, along all software packages despite technical and functional variabilities among the set of software packages. Not all qualities are quantitatively measured. For example, performance is not measured by running the code. A surface investigation of the documentation of each software package for performance information is conducted instead. The measured software packages are open source, and it is recommended to keep this in mind when considering the recommendations of this report. Some recommendations may not apply to closed-source software. Practices surrounding close source software development may differ.

@@

\begin{itemize}
	\item \textbf{Introduction} to the report, including its purpose, motivation, scope, and organization.
	\item LBM software \textbf{Domain Analysis}. A domain analysis consists of systematically identifying and documenting the commonalities, variabilities, and terminology of a software family \citep{Weiss1997}.
	The purpose of this analysis in this state of the practice assessment is to better classify the different software products based on their functionality.
	\item \textbf{Methodology} of this state of the practice assessment, including the steps of the methodology, software quality definitions and how these qualities are assessed in this report, an overview of candidate software selection, empirical measures gathered, and the analytical hierarchy process. This follows a general methodology for assessing the state of the practice for SCS domains.
	\item Quantitative and qualitative \textbf{Results}. 
	\item An analysis of the results and \textbf{Answers To Research Questions}.
	\item \textbf{Conclusion} to the report including comments on suggestions for future state of the practice assessments.
	\item \textbf{Appendix}: This section includes our Research Questions, the Measurement Template, Grading Template, Ethics Approval, and Developer Interview Questions.
\end{itemize}

@@

We added this domain commonality and variability analysis to our original methodology because the first time we did the State of the Practice exercise we found we weren't comparing "apples to apples". Software packages can have considerable variation, even when the packages appear to be in the same software family. Viewing the ranking exercise as a means for selecting the right tool for a job, functionality is important. A 2D Lattice Boltzmann solver is not a substitute for a 3D Lattice Boltzmann solver when the latter is required. Therefore, this exercise seeks to classify the different software products based on their functionality. Nonfunctional qualities are assessed through other means in our process. For instance, through usability experiments as noted in the \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Methodology.pdf}{Methodology for Assessing the State of the Practice for Domain X} document.

@@

LBM are a family of fluid dynamics algorithms for simulating single-phase and multiphase fluid flows, often incorporating additional physical complexities \citep{chen1998lattice}. They consider the behaviours of a collection of particles as a single unit at the mesoscopic scale, between the nanoscopic and microscopic scales. These methods predict the positional probability of a collection of particles moving through a lattice structure. Off-the-shelf (OTS) Lattice Boltzmann Solvers (LBS) allow for a range of fluid and physical model input parameters, computational parameters, and output parameters.

@@

This section lists common features among potential family members. The commonalities are organized using the following abstraction of the system, which can be used to describe all Lattice Boltzmann systems: input information, generate the simulation, output the results. Section \ref{comm_lbm} describes the commonalities for the simulation step. Section \ref{comm_in} highlights the input information that is required for Lattice Boltzmann systems. The next section, Section \ref{comm_out}, shows the common features for the output of Lattice Boltzmann systems, such as the requirement that mesh information be written to files. Although the output information could simply be written to the computer’s memory, in all practical applications it is desirable to have a persistent record of the output that was created. The final section covers nonfunctional requirements of the system. For instance, all systems will have the goal that the response time to a user’s request is small enough to allow the user to focus on his/her problem and to maintain his/her train of thought, without excessive waiting time. The commonality in this case is refined by a later variability because the specific requirement on the response time will depend on the individual software packages, their simulation models, inputs, and outputs.

@@

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & Every Lattice Boltzmann version uses a Boltzmann transport equation to describe the statistical behaviour of a system that does not have collisions. The equation is $(\frac{d}{dt} + \mathrm{e} \cdot \nabla_\mathrm{x} + \frac{F}{m} \cdot \nabla_\mathrm{e})f = \Omega(t)$, where $\mathrm{e}$ is the velocity, $\nabla_\mathrm{x}$ is the position vector gradient, $\nabla_\mathrm{e}$ is the velocity gradient, and $\Omega(t)$ is the collision operator as a function of time.\\
		\hline
		\bf Related Variability & \vref{vartraneq}\\
		\hline
	\end{tabular}
\end{minipage}\\
\subsubsection{Input}\label{comm_in}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum\label{fluidandboundaryinfo} \\
		\hline
	\end{tabular}\\

@@

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & LBS write fluid predictions, such as the output of the transport equation, to memory.\\
		\hline
		\bf Related Variability & \vref{varvisual} \vref{formatoutput} \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Variabilities}\label{var_sec}
This section provides a list of characteristics that may vary among family members. As in Section \ref{comm_sec}, the first three subsections on variabilities are organized into the following sublists: Simulation Models, Input and Output. The final two subsections list variabilities that can be characterized as system constraints and as nonfunctional requirements.

@@

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & The number of fluids allowed in the LBS.\\
		\hline
		\bf Related Commonality & \cref{fluidandboundaryinfo}\\
		\hline
		\bf Related Parameter & \pref{parfluids}\\
		\hline
	\end{tabular}
\end{minipage}
~\newline
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number&  V\refstepcounter{varnum}\thevarnum\label{typefluidparm} \\
		\hline
	\end{tabular}\\

@@

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & Amount of storage and memory needed for the LBS.\\
		\hline
		\bf Related Commonality & None \\
		\hline
		\bf Related Parameter & \pref{parstorageconst}\\
		\hline
	\end{tabular}
\end{minipage}\\
\newpage
\subsection{Parameters of Variation}
This section specifies the parameters of variation for the variabilities listed in Section \ref{var_sec}. They are organized into the same five subcategories as employed previously: Simulation Models, Input, Output, System Constraints, Nonfunctional Requirements.

Each parameter of variation is given a unique identifier of the form “P” followed by a natural number. The corresponding variability is listed and a hyperlink is provided that allows navigation back to the appropriate item in Section \ref{var_sec}. The final entry for each parameter of variation is the binding time, which is the time in the software lifecycle when the variability is fixed. The binding time could be during specification, or during building
of the system (compile time), or during execution of the system (run time). It is possible to have a mixture of binding times. For instance, a parameter of variation could have a binding time of “specification or building” to represent that the parameter could be set at specification time, or it could be postponed until the given family member is built. The choice of postponing the decision until the build could be associated with the presence of a domain specific language that would allow postponing decisions on the values of the parameter of variation.

@@

In this project we set out to answer the research questions listed in Section \ref{purpose} for the LBM SCS domain. The process involved systematically measuring and analyzing members of this software family along the quality attributes found in Section \ref{softwarequalities}. This included gathering quantitative and qualitative measurements. A goal of the project was to produce a quality assessment for LBM software packages.

The methodology used in this LBM software assessment is a general state of the practice methodology that can be applied to any SCS domain. It was developed as an update to previous state of the practice exercises. This general methodology is linked in the next subsection.

We collected quantitative data using the measures found in the measurement template of Section \ref{measurementtemplate}. Some of this was empirical software engineering related data, such as the number of files, number of lines of code (LOC), percentage of issues that are closed, etc. Most of the data was gathered by manually investigating the software, its source code, and its artifacts, while some was gathered using the empirical measurement tools discussed in Section \ref{empiricalmeasures}.

We also collected qualitative data by interviewing the software package developers and asking them the questions found in the Appendix in \ref{interviewquestions}. Furthermore, we solicited the assistance of domain experts to better assess each software package by leveraging their experience to assess the functional and non-functional requirements for the software domain.

This section begins by describing the steps of the overall process we used to select, measure and compare LBM software. This is followed by quality definitions and how these qualities were assessed in our project. The rest of the section provides an overview of how candidate software packages were selected and filtered, the empirical measurements and software tools that were used, and the analytic hierarchy process (AHP).

@@

\begin{enumerate}
	\item List candidate software packages for the domain. This is discussed in Section \ref{identifysoftware}.
	\item Filter the software package list. This is discussed in Section \ref{filtersoftware}.
	\item Gather the source code and documentation for each software package.
	\item Collect empirical measures. This is discussed in Section \ref{empiricalmeasures}.
	\item Measure using the measurement template. This is discussed in Section \ref{empiricalmeasures}. The measurement template can be found in the Appendix in \ref{measurementtemplate2}.
	\item Survey the developers. The developers of each software package in the filtered software list were contacted for voluntary interviews. The interview questions can be found in the Appendix in \ref{interviewquestions}.
	\item Use AHP to rank the software packages. This is discussed in Section \ref{AHP}.
	\item Analyze the results and answer the research questions. The answers can be found in Section \ref{answersquestions}.
\end{enumerate}

@@

All of the quality measurements that are part of the AHP analysis are empirical measurements. Qualitative data gathered during interviews with developers is not part of the AHP analysis. This part of the assessment focuses on data that is reasonably easy to collect. Much of the data was gathered by manually reviewing the artifacts of each software package, while some was gathered using the freeware tools discussed below. The data that was gathered is listed in the measurement template found in the Appendix in \ref{measurementtemplate2}. Some of the data required processing other data within the template, including the status of the software package, which relied on the last commit date; the percentage of issues that are closed, which relied on the number of open and closed issues; and the percentage of code that is comments, which relied on the number of total lines and comment lines in text-based files. The complete measurement template data was then analyzed using the grading template found in the Appendix in \ref{gradingtemplate2}, the output of which was analyzed using the AHP described in Section \ref{AHP}.

@@

The Analytical Hierarchy Process (AHP) is a decision-making technique that is used to compare multiple options by multiple criteria. In our work AHP was used for comparing and ranking the LBM software packages using the overall impression quality scores that were gathered in the measurement template found in the Appendix in \ref{measurementtemplate2} using the grading template found in the Appendix in \ref{gradingtemplate2}. AHP performs a pairwise analysis using a matrix and generates an overall score as well as individual quality scores for each software package. \cite{SmithEtAl2016} shows how AHP is applied to ranking software based on quality measures.

@@

Figure \ref{Fig_Robustness} shows the surface robustness ranking of the software packages using AHP. Software packages with a high score behaved reasonably in response to unexpected input as described above. All of the software packages that installed correctly passed this test. They output descriptive error messages or did not crash. Software packages with a lower surface robustness score had not installed correctly, so their robustness score may not be a true reflection of runtime robustness. Similarly, all software packages that installed correctly and require plain text input files correctly handled an unexpected change to these input files, including a replacement of new lines with carriage returns. Furthermore, nine of the top 10 ranked packages are noted as being alive.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user manual, documented user characteristics, and a user support model. In total 18 software packages have a tutorial and 13 have a user manual. The tutorials vary in scope and substance, and eight include an expected output. Most user manuals are in the form of a file that can be downloaded, while some are rendered on a web-page. Some packages (waLBerla) do not have a user manual, but do have useful documentation distributed on their web-pages. Expected user characteristics are documented in four software packages (laboetie, LIMBES, Ludwig, Palabos). Users are typically scientists or engineers. Their background is often physics, chemistry, biophysics, or mathematics. All but one of the packages (LIMBES) have a user support model, and many of them have multiple avenues of user support. The most popular avenue of support is Git, followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Issue tracking is used in 22 software packages, 15 of which use Git, six use email, and one (SunlightLB) uses SourceForge. Most software packages that use Git have most of their issues closed, and only three (laboetie, lettuce, Sailfish) have less than 50 percent of their issues closed. Alive packages (11 use Git issue tracking) have 64\% of their issues closed, while dead packages (3 use Git issue tracking) have 71\% of their issues closed. This information is presented in Table \ref{gitrepodata}. Furthermore, 13 packages that use Git for issue tracking use GitHub as a version control system, while two (Palabos, waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS for issue tracking, and seven packages do not appear to use any issue tracking system.

Software package code files were further measured for the percentage of code that is comments. The findings are presented in Table \ref{gitrepodata}. Packages with a higher percentage of comments were designated as more maintainable. Comments represent more than 10 percent of code files in 15 packages, and the average percentage of code comments is about 14.\\

@@

Figure \ref{Fig_Maintainability} shows the maintainability ranking of the software packages using AHP. Software packages with a high score provide version numbers on documents and source code releases, have an abundance of high quality artifacts, and use an issue tracking tool and version control system. These packages also appear to reasonably handle tracked issues, having most of their issues closed. Their code files are well commented with more than 10 percent of the code being comments. Furthermore, nine of the top 10 ranked packages are noted as being alive.

@@

Figure \ref{Fig_Reusabilty} shows the reusability ranking of the software packages using AHP. Software packages with a high score have thousands of source code files and API documentation. The highest scoring packages, ESPResSo and waLBerla, have extensive functionality, including graphical visualizations as well as modeling that does not use LBM. For this reason a comparison with other software packages is not on a level field. However, these packages do have an abundance of reusable components. Furthermore, nine of the top 10 ranked packages are noted as being alive.

Table \ref{moduledata} shows file and line of code data of the software packages. Packages with a high reusability score do not have many LOC per text file, generally having a few hundred lines or less. This suggests that the source code of these packages is likely functionally modularized, and modules could be reused in other projects.\\

@@

@@

Ten random source code files of each software package were reviewed for several measures. This assessment of surface understandability may not perfectly reflect each package due to the limitation of only testing 10 files.

>>>>

\documentclass[12pt, notitlepage]{article}
\usepackage{amsmath, mathtools}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{xr}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{xfrac}
\usepackage{tabularx}
\usepackage{float}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{titling}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage{indentfirst}
\usepackage[toc,page]{appendix}

@@

We analyze the state of the practice of software development of Lattice Boltzmann Methods software domain by quantitatively and qualitatively measuring, and comparing, 23 software packages. A general methodology for assessing the state of the practice of software development in scientific computing software domains is presented. A domain analysis of the Lattice Boltzmann Methods software family is made, and candidate software packages are identified and curated. The packages are assessed to answer software development related research questions to understand how software quality is impacted by software development choices, including principles, processes, and tools. Software developers are interviewed to identify development pain points, and to identify how software quality is ensured. Quantitative data is used to rank the software packages using the Analytical Hierarchy Process. The packages Ludwig, ESPResSo, and Palabos are ranked as the top three packages. Each of them scored high in at least several of the individual qualities that were quantitatively measured. The ranking designations are compared with rankings from the software development community. Recommendations for improving software along quality metrics are made. These include providing a detailed user manual and tutorials, explicitly stating the limits of the software, using user-friendly software languages, considering a peer review process, communicating development and contribution information, and using continuous integration and project management tools. This knowledge can be used to guide future development of scientific computing software along quality attributes to reduce software quality failures.\\

@@

We analyze the development of Computational Fluid Dynamics (CFD) software packages that use Lattice Boltzmann Methods (LBM). These LBM algorithms consider the behaviours of a collection of particles as a single unit of an intermediate size, and predict their positional probability as they move through a lattice structure. Figure \ref{circularflow} presents the streamlines of converged flow past a stationary circular cylinder with varied Reynolds number, a common task of LBM.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{circularflow}
		\caption{Streamlines of flow past a stationary circular cylinder at Re = 10, 20, and 40 \citep{chen2021phase}}
		\label{circularflow}
	\end{center}
\end{figure}

@@

The packages are assessed for a given set of quality attributes using quantitative and qualitative data. The qualities are defined in Section \ref{softwarequalities}. Quantitative data is measured using quality metrics. The findings are presented in Section \ref{AHPresults}. Qualitative data is gathered from interviews with software package developers. The interview findings are presented in Section \ref{interviewresults}. In this context the purposes of the report are as follows:

@@

In the course of this assessment we updated the methodology that was used during previous assessments. Details of the new methodology are in \citep{methodology2020}. The previous set of research questions were critically assessed and modified. The updated questions are listed in Section \ref{purpose}. In this re-boot we collected more quantitative and qualitative data, focused on software quality attributes, and added more measures, including collecting empirical data and interviewing developers. We have also added a domain analysis to better characterize the functionality provided by the software, and have leveraged the expertise of a domain expert. As in past assessments, the collected data was combined to rank the software using the Analytical Hierarchy Process (AHP). The domain expert was consulted to verify the ordering.

@@

We analyze a filtered set of LBM software packages along quantitative and qualitative measures. Many of these measures are captured using surface measurements, which can be categorized as initial and easy to capture measurements, of the underlying quality; they may not represent the true quality of the software. Surface measurements are taken as they allow us to apply the same measurement, with reasonable effort, along all software packages despite technical and functional variabilities among the set of software packages. Not all qualities are quantitatively measured. For example, performance is not measured by running the code. A surface investigation of the documentation of each software package for performance information is conducted instead. The measured software packages are open source. Some recommendations may not apply to closed-source software. Practices surrounding close source software development may differ.

@@

\begin{itemize}
	\item \textbf{Introduction} to the report, including its purpose, motivation, scope, and organization.
	\item LBM software \textbf{Domain Analysis}. A domain analysis consists of systematically identifying and documenting the commonalities, variabilities, and terminology of a software family \citep{Weiss1997}.
	The purpose of this analysis in this state of the practice assessment is to better classify the different software products based on their functionality.
	\item \textbf{Methodology} of this state of the practice assessment, including the steps of the methodology, software quality definitions and how these qualities are assessed in this report, an overview of candidate software selection, empirical measures gathered, and the AHP. This follows a general methodology for assessing the state of the practice for SCS domains.
	\item Quantitative and qualitative \textbf{Results}. 
	\item An analysis of the results and \textbf{Answers To Research Questions}.
	\item \textbf{Conclusion} to the report including comments on suggestions for future state of the practice assessments.
	\item \textbf{Appendix}: This section includes our Research Questions, the Measurement Template, Grading Template, Ethics Approval, and Developer Interview Questions.
\end{itemize}

@@

We added this domain commonality and variability analysis to our original methodology because the first time we did the State of the Practice exercise we found we weren't comparing ``apples to apples". Software packages can have considerable variation, even when the packages appear to be in the same software family. Viewing the ranking exercise as a means for selecting the right tool for a job, functionality is important. A 2D Lattice Boltzmann solver is not a substitute for a 3D Lattice Boltzmann solver when the latter is required. Therefore, this exercise seeks to classify the different software products based on their functionality. Nonfunctional qualities are assessed through other means in our process. For instance, through usability experiments as noted in the \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Methodology.pdf}{Methodology for Assessing the State of the Practice for Domain X} document.

@@

LBM are a family of fluid dynamics algorithms for simulating single-phase and multiphase fluid flows, often incorporating additional physical complexities \citep{chen1998lattice} such as reflective and non-reflective boundaries. They consider the behaviours of a collection of particles as a single unit at the mesoscopic scale, between the nanoscopic and microscopic scales. These methods predict the positional probability of a collection of particles moving through a lattice structure. Off-the-shelf (OTS) Lattice Boltzmann Solvers (LBS) allow for a range of fluid and physical model input parameters, computational parameters, and output parameters.

@@

This section lists common features among potential family members. The commonalities are organized using the following abstraction of the system, which can be used to describe all Lattice Boltzmann systems: input information, generate the simulation, output the results. Section \ref{comm_lbm} describes the commonalities for the simulation step. Section \ref{comm_in} highlights the input information that is required for Lattice Boltzmann systems. The next section, Section \ref{comm_out}, shows the common features for the output of Lattice Boltzmann systems, such as the requirement that mesh information be written to files. Although the output information could simply be written to the computer’s memory, in all practical applications it is desirable to have a persistent record of the output that was created.

@@

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & Every Lattice Boltzmann version uses a Boltzmann transport equation to describe the statistical behaviour of a system that does not have collisions. The equation is $(\frac{d}{dt} + \mathrm{e} \cdot \nabla_\mathrm{x} + \frac{F}{m} \cdot \nabla_\mathrm{e})f = \Omega(t)$, where $\mathrm{e}$ is the velocity, $\nabla_\mathrm{x}$ is the position vector gradient, $F$ is the force of the fluid, $m$ is the mass, $\nabla_\mathrm{e}$ is the velocity gradient, $f$ indicates the probability density function, and $\Omega(t)$ is the collision operator as a function of time.\\
		\hline
		\bf Related Variability & \vref{vartraneq}\\
		\hline
	\end{tabular}
\end{minipage}\\
\subsubsection{Input}\label{comm_in}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum\label{fluidandboundaryinfo} \\
		\hline
	\end{tabular}\\

@@

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & LBS write fluid predictions, such as the output of the transport equation, to memory.\\
		\hline
		\bf Related Variability & \vref{varvisual} \vref{formatoutput} \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Variabilities}\label{var_sec}
This section provides a list of characteristics that may vary among family members. As in Section \ref{comm_sec}, the first three subsections on variabilities are organized into the following sublists: Simulation Models, Input and Output. The final subsection lists variabilities that can be characterized as system constraints.

@@

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & The number of fluids allowed in a multifluid simulation.\\
		\hline
		\bf Related Commonality & \cref{fluidandboundaryinfo}\\
		\hline
		\bf Related Parameter & \pref{parfluids}\\
		\hline
	\end{tabular}
\end{minipage}
~\newline
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number&  V\refstepcounter{varnum}\thevarnum\label{typefluidparm} \\
		\hline
	\end{tabular}\\

@@

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & Amount of storage and memory needed for the LBS.\\
		\hline
		\bf Related Commonality & None \\
		\hline
		\bf Related Parameter & \pref{parstorageconst}\\
		\hline
	\end{tabular}
\end{minipage}\\
\newpage
\subsection{Parameters of Variation}
This section specifies the parameters of variation for the variabilities listed in Section \ref{var_sec}. They are organized into the same five subcategories as employed previously: Simulation Models, Input, Output, and System Constraints.

Each parameter of variation is given a unique identifier of the form “P” followed by a natural number. The corresponding variability is listed and a hyperlink is provided that allows navigation back to the appropriate item in Section \ref{var_sec}. The final entry for each parameter of variation is the binding time, which is the time in the software lifecycle when the variability is fixed. The binding time could be during specification, or during building
of the system (build time), or during execution of the system (run time). It is possible to have a mixture of binding times. For instance, a parameter of variation could have a binding time of “specification or building” to represent that the parameter could be set at specification time, or it could be postponed until the given family member is built. The choice of postponing the decision until the build could be associated with the presence of a domain specific language that would allow postponing decisions on the values of the parameter of variation.

@@

In this project we set out to answer the research questions listed in Section \ref{purpose} for the LBM SCS domain. The process involved systematically measuring and analyzing members of this software family along the quality attributes listed in the research questions and described in Section \ref{softwarequalities}. The methodology includes gathering quantitative and qualitative measurements. A goal of the project is to produce a quality assessment for LBM software packages.

The methodology used in this LBM software assessment is a general state of the practice methodology that can be applied to any SCS domain. It was developed as an update to previous state of the practice exercises.

We collected quantitative data using the measures found in the measurement template in Appendix \ref{measurementtemplate2}. Some of this was empirical software engineering related data, such as the number of files, number of lines of code (LOC), percentage of issues that are closed, etc. Most of the data was gathered by manually investigating the software, its source code, and its artifacts, while some was gathered using the empirical measurement tools discussed in Section \ref{empiricalmeasures}.

We also collected qualitative data by interviewing the software package developers and asking them the questions found in Appendix \ref{interviewquestions}. Ethics clearance information can be found in Appendix \ref{ethicsapproval}. Furthermore, we solicited the assistance of domain experts to better assess each software package by leveraging their experience to assess the functional and non-functional requirements for the software domain.

This section begins by describing the steps of the overall process we used to select, measure and compare LBM software. This is followed by quality definitions and how these qualities were assessed in our project. The rest of the section provides an overview of how candidate software packages were selected and filtered, the empirical measurements and software tools that were used, and the AHP.

@@

\begin{enumerate}
	\item List candidate software packages for the domain. This is discussed in Section \ref{identifysoftware}.
	\item Filter the software package list. This is discussed in Section \ref{filtersoftware}.
	\item Gather the source code and documentation for each software package.
	\item Collect empirical measures. This is discussed in Section \ref{empiricalmeasures}.
	\item Measure using the measurement template. This is discussed in Section \ref{empiricalmeasures}. The measurement template can be found in Appendix \ref{measurementtemplate2}.
	\item Survey the developers. The developers of each software package in the filtered software list were contacted for voluntary interviews. The interview questions can be found in Appendix \ref{interviewquestions}.
	\item Use AHP to rank the software packages. This is discussed in Section \ref{AHP}.
	\item Analyze the results and answer the research questions. The answers can be found in Section \ref{answersquestions}.
\end{enumerate}

@@

All of the quality measurements that are part of the AHP analysis are empirical measurements. Qualitative data gathered during interviews with developers is not part of the AHP analysis. This part of the assessment focuses on data that is reasonably easy to collect. Much of the data was gathered by manually reviewing the artifacts of each software package, while some was gathered using the freeware tools discussed below. The data that was gathered is listed in the measurement template found in Appendix \ref{measurementtemplate2}. Some of the data required processing other data within the template, including the status of the software package, which relied on the last commit date; the percentage of issues that are closed, which relied on the number of open and closed issues; and the percentage of code that is comments, which relied on the number of total lines and comment lines in text-based files. The complete measurement template data was then analyzed using the grading template found in Appendix \ref{gradingtemplate2}, the output of which was analyzed using the AHP described in Section \ref{AHP}.

@@

The Analytical Hierarchy Process (AHP) is a decision-making technique that is used to compare multiple options by multiple criteria. In our work AHP was used for comparing and ranking the LBM software packages using the overall impression quality scores that were gathered in the measurement template found in Appendix \ref{measurementtemplate2} using the grading template found in Appendix \ref{gradingtemplate2}. AHP performs a pairwise analysis using a matrix and generates an overall score as well as individual quality scores for each software package. Smith et al. (2016) shows how the AHP is applied to ranking software based on quality measures \cite{SmithEtAl2016}.

@@

Figure \ref{Fig_Robustness} shows the surface robustness ranking of the software packages using AHP. Software packages with a high score behaved reasonably in response to unexpected input as described above. All of the software packages that installed correctly passed this test. They output descriptive error messages or did not crash. Software packages with a lower surface robustness score had not installed correctly, so their robustness score may not be a true reflection of runtime robustness. Similarly, all software packages that installed correctly and require plain text input files correctly handled an unexpected change to these input files, including a replacement of new lines with carriage returns. Furthermore, nine of the top 10 ranked packages are noted as being alive. LIMBES is noted as being dead.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user manual, documented user characteristics, and a user support model. In total 18 software packages have a tutorial, 13 have a user manual, and 11 have both. The tutorials vary in scope and substance, and eight include an expected output. Most user manuals are in the form of a file that can be downloaded, while some are rendered on a web-page. Some packages (waLBerla) do not have a user manual, but do have useful documentation distributed on their web-pages. Expected user characteristics are documented in four software packages (laboetie, LIMBES, Ludwig, Palabos). Users are typically scientists or engineers. Their background is often physics, chemistry, biophysics, or mathematics. All but one of the packages (LIMBES) have a user support model, and many of them have multiple avenues of user support. The most popular avenue of support is Git, followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Issue tracking is used in 22 software packages, 15 of which use Git, six use email, and one (SunlightLB) uses SourceForge. Most software packages that use Git have most of their issues closed, and only three (laboetie, lettuce, Sailfish) have less than 50 percent of their issues closed. Four of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, LUMA) have most of their issues closed. Fourth ranked LUMA does not use Git. Alive packages (11 use Git issue tracking) have 64\% of their issues closed, while dead packages (3 use Git issue tracking) have 71\% of their issues closed. This information is presented in Table \ref{gitrepodata}. Furthermore, 13 packages that use Git for issue tracking use GitHub as a version control system, while two (Palabos, waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS for issue tracking, and seven packages do not appear to use any issue tracking system.

Software package code files were further measured for the percentage of code that is comments. The findings are presented in Table \ref{gitrepodata}. Packages with a higher percentage of comments were designated as more maintainable. Comments represent more than 10 percent of code files in 15 packages, and the average percentage of code comments is about 14 percent. Four of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, OpenLB) have more than the average. Fifth ranked LUMA has only 0.2 percent comments, the fewest of any package. The package has the most lines of source code, with over four million. The next largest package is ESPResSo++ with one million.

Figure \ref{Fig_Maintainability} shows the maintainability ranking of the software packages using AHP. Software packages with a high score provide version numbers on documents and source code releases, have an abundance of high quality artifacts, and use an issue tracking tool and version control system. These packages also appear to reasonably handle tracked issues, having most of their issues closed. Their code files are well commented with more than 10 percent of the code being comments. Furthermore, nine of the top 10 ranked packages are noted as being alive. MP-LABS is noted as being dead.\\

@@

@@

Figure \ref{Fig_Reusabilty} shows the reusability ranking of the software packages using AHP. Software packages with a high score have thousands of source code files and API documentation. The highest scoring packages, ESPResSo and waLBerla, have extensive functionality, including graphical visualizations as well as modeling that does not use LBM. For this reason a comparison with other software packages is not on a level field. However, these packages do have an abundance of reusable components. Furthermore, nine of the top 10 ranked packages are noted as being alive. HemeLB is noted as being dead.

@@

Table \ref{moduledata} shows file and line of code data of the software packages. Packages with a high reusability score do not have many LOC per text file, generally having a few hundred lines or less. This suggests that the source code of these packages is likely functionally modularized, and modules could be reused in other projects.

@@

Ten random source code files of each software package were reviewed for several measures. This assessment of surface understandability may not perfectly reflect each package due to the practical limitation of only being able to test 10 files.

All of the packages appear to have a consistent indentation and formatting style. Only LUMA and HemeLB explicitly identify coding standards that are used during development. Generally, the software packages use consistent, distinctive, and meaningful code identifiers. Only four packages (LB2D-Prime, LB3D-Prime, LIMBES, MP-LABS) appear to use vague identifiers, such as single letters for variables. Symbolic constants were observed in the source code of 12 packages. The constants are used for various parameters, mathematical constants, and matrix definitions. All of the packages are well commented and the comments clearly indicate what is being done. Domain algorithms are noted in the source code of 11 packages. Table \ref{moduledata} suggests that the software packages are modularized to various degrees. When observing the source code files, it was found that 13 of the packages have a consistent style and order of function parameters.
